Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-10 ind_files/04_unification_discussion.tex
Minor phrasing changes

To the best of my knowledge, a general formulation of stochastic field equations for the population dynamics of quantitative traits from the functional Fokker-Planck side in the manner I have carried out here has not been done before. Similar equations have been formulated for some specific stochastic models of quantitative trait evolution~\citep{rogers_demographic_2012,rogers_modes_2015} and population ecology of size-structured communities~\citep{odwyer_integrative_2009}. Stochastic field equations are also known in mathematical neurobiology~\citep{buice_field-theoretic_2007,bressloff_stochastic_2010,coombes_neural_2014}, and have recently been proposed in a model of collective motion~\citep{o_laighleis_minimal_2018}, suggesting that the study and formalization of such equations by pure mathematicians could have wide-spread applications. As is, such field equations are not on very solid mathematical footing, and are currently primarily used by physicists working in areas such as statistical field theory and are largely attacked using ingenious (and often heuristic) arguments and tools that are not necessarily mathematically well understood~\citep{carmona_stochastic_1999}. Equation \eqref{disc_functional_field_eqns} opens up the study of quantitative trait dynamics in finite fluctuating populations to analysis using some of these tools --- For example, a careful reading of the literature provides many hints for how the major ideas behind the Fock space representation~\citep{del_razo_probabilistic_2022}, perturbative `loop diagrams/expansions'~\citep{dodd_many-body_2009}, and the path integral (really a `functional integral') formalism~\citep{doi_second_1976, peliti_path_1985, chow_path_2015, weber_master_2017} can all be co-opted to attack the sort of stochastic processes I formulate as a model for evolutionary dynamics via equation \eqref{disc_functional_field_eqns}. Equation \eqref{disc_functional_field_eqns} also provides a more accessible alternative formulation that does not (explicitly) rely on tools from relatively `advanced' mathematical fields such as martingale theory and measure-valued processes that are typically used to study the evolution of quantitative traits in finite, fluctuating populations~\citep{dawson_stochastic_1975,fleming_measure-valued_1979,ethier_markov_1986,champagnat_unifying_2006,etheridge_mathematical_2011,week_white_2021}. This is intended to make the models and ideas more accessible to theorists without formal training in mathematical areas such as measure theory and functional analysis.

>>>>

To the best of my knowledge, a general formulation of stochastic field equations for the population dynamics of quantitative traits from the functional Fokker-Planck side in the manner I have carried out here has not been done before. Similar equations have been formulated for some specific stochastic models of quantitative trait evolution~\citep{rogers_demographic_2012,rogers_modes_2015} and population ecology of size-structured communities~\citep{odwyer_integrative_2009}. Stochastic field equations are also known in mathematical neurobiology~\citep{buice_field-theoretic_2007,bressloff_stochastic_2010,coombes_neural_2014}, and have recently been proposed in a model of collective motion~\citep{o_laighleis_minimal_2018}, suggesting that the study and formalization of such equations by pure mathematicians could have wide-spread applications. Such field equations are currently primarily used by physicists working in areas such as statistical field theory or quantum field theory and are largely attacked using ingenious (and often heuristic) arguments and tools that may not necessarily be mathematically well understood~\citep{carmona_stochastic_1999}. Equation \eqref{disc_functional_field_eqns} opens up the study of quantitative trait dynamics in finite fluctuating populations to analysis using some of these tools --- For example, a careful reading of the literature provides many hints for how the major ideas behind the Fock space representation~\citep{del_razo_probabilistic_2022}, perturbative `loop diagrams/expansions'~\citep{dodd_many-body_2009}, and the path integral formalism~\citep{doi_second_1976, peliti_path_1985, chow_path_2015, weber_master_2017} can all be co-opted to coax biological insights from the sort of stochastic processes I formulate as a model for evolutionary dynamics via equation \eqref{disc_functional_field_eqns}. Equation \eqref{disc_functional_field_eqns} also provides a more accessible alternative formulation that does not (explicitly) rely on tools from relatively `advanced' mathematical fields such as martingale theory and stochastic analysis that are typically used to study the evolution of quantitative traits in finite, fluctuating populations~\citep{dawson_stochastic_1975,fleming_measure-valued_1979,ethier_markov_1986,champagnat_unifying_2006,etheridge_mathematical_2011,week_white_2021}. This is intended to make the models and ideas more accessible to theorists without formal training in mathematical areas such as measure theory and functional analysis.
